---
title: "Community analysis"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="./figs/")
```
```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(vegan)
library(indicspecies)
library(HierDpart)
library(leaflet)
source("../R/iDIP.R")
```

```{r getData}
combdat<-read.csv("cleanedCombinedCommunity.csv",stringsAsFactors = TRUE)
levels(combdat$SeagrassCover)<-c("bare","sparse","patchy","dense")
```
```{r}
seasonCols<-c(Spring="coral",Summer="olivedrab3",Fall="gold",Winter="dodgerblue")
seagrassCs<-c(bare="coral",sparse="olivedrab3",patchy="gold",dense="dodgerblue")#c(bare="#ffffcc",sparse="#c2e699",patchy="#78c679",dense="#238443")
seasonPchs<-c(Spring=15,Summer=17,Fall=18,Winter=19)
seagrassPs<-c(bare=16, sparse=18, patchy=17, dense=15)
```


## Community analysis


### dbrda


```{r}
abunds<-as.matrix(combdat[rowSums(combdat[,6:ncol(combdat)])>4,6:ncol(combdat)])
ys<-combdat[rowSums(combdat[,6:ncol(combdat)])>4,c("SeagrassCover","Season","TowDist")]


seasonRDA<-capscale(abunds~SeagrassCover*Season + Condition(TowDist),
                 ys,
                 dist="bray")
```
```{r}
plot(seasonRDA)
anova(seasonRDA,by="terms")
seasonSum<-summary(seasonRDA)
```

These results suggest that season and seagrass cover -- and their interaction -- are important in determining the community in seagrass beds. 




```{r RDAplot, fig.cap="dbRDA plot of the first two axes, which describe >50% of the constrained variation. The colours of points represent seagrass cover categories (bare, sparse, patchy, dense) and the shapes of the points represent the season during which the sample was collected (winter, spring, summer, fall). Overlaid on top are the species, with flounder and shrimp being the most obvious species driving divergence."}

plot(seasonRDA,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(seasonSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(seasonSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(seasonSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(seasonSum$concont$importance[2,2]*100,2),
       "% of constrained)")
     )
points(seasonRDA,
       cex=1.5,
       col=scales::alpha(
         seagrassCs[ys$SeagrassCover],0.5
       ),
       pch=seasonPchs[ys$Season])
text(seasonRDA, scaling = 2, display = "sp") 
```

### Indicator species analysis

```{r}
coverIndics<-multipatt(abunds,ys$SeagrassCover)
summary(coverIndics)
```

```{r}
seasonIndics<-multipatt(abunds,ys$Season)
summary(seasonIndics)
```


#### exploring indicators

```{r getGPS}
gpsdat<-read.csv("../data/gpsdat.csv")
```

```{r}
lat <- c(-43.794,-43.752) #define our map's ylim
lon <- c(172.930,172.933) #define our map's xlim
center <- c(mean(c(gpsdat$lat_start,gpsdat$lat_stop)), 
            mean(c(gpsdat$long_start,gpsdat$long_stop)))  #tell what point to center on
zoom <- 16  #zoom: 1 = furthest out (entire globe), larger numbers = closer in

```


```{r seasonIndics}
kahawai<-combdat[combdat$kahawai>0,]
limpets<-combdat[combdat$limpets>0,] # not abundant enough
smelt<-combdat[combdat$smelt>0,] # lots of smelt on one day
blueCod<-combdat[combdat$blue_cod_juv>0,] # almost all from one day
```{r makeSpeciesMap}
make_species_map<-function(gpsdat,speciesdat,speciesName,
                           center,zoom,color="yellow",
                           scale=5){
  long<-rowMeans(gpsdat[gpsdat$TowNum %in% speciesdat$TowNum,
                       c("long_start","long_stop")])
  lat<-rowMeans(gpsdat[gpsdat$TowNum %in% speciesdat$TowNum,
                       c("lat_start","lat_stop")])
  radius<-speciesdat[,speciesName]
  
  m <- leaflet() %>%
    addTiles() %>%  # Add default OpenStreetMap map tiles
    setView(lng=center[2], lat=center[1],zoom=zoom) %>% 
     #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
    addProviderTiles("Esri.WorldImagery") %>%
    addCircles(lng = long, lat = lat, weight = 1,
      radius = radius*scale,col=color)
  
  m
  return(m)
}

```

```{r}
long<-rowMeans(gpsdat[gpsdat$TowNum %in% kahawai$TowNum,
                      c("long_start","long_stop")])
lat<-rowMeans(gpsdat[gpsdat$TowNum %in% kahawai$TowNum,
                     c("lat_start","lat_stop")])
radius<-kahawai$kahawai

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(lng=center[2], lat=center[1],zoom=zoom) %>% 
   #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
  addProviderTiles("Esri.WorldImagery") %>%
  addCircles(lng = long, lat = lat, weight = 1,
    radius = radius*5,col="yellow")

m




```


```{r seagrassIndics}
# not bare
arrowCrab<-combdat[combdat$arrowheadCrab>0,]
blenny<-combdat[combdat$blenny>0,]
stigmatopora<-combdat[combdat$s.nigra>0,]
mudWhelk<-combdat[combdat$mud_whelk>0,]

#sparse: also kahawai
spotty<-combdat[combdat$spotty>0,] 
bristleworm<-combdat[combdat$bristleworm>0,] 
#patchy: smelt
```



## Diversity analysis

### Rarefaction

```{r}
rar<-rrarefy(abunds)
```


### iDIP analysis


```{r iDIP}
stru<-Str(n=nrow(abunds),
          nreg=length(unique(combdat$Date)),
          r=tapply(combdat$SeagrassCover,combdat$Date,length))

stru<-t(data.frame(Island=rep("South Island",ncol(abund)),
                   Estuary=gsub("(\\w+)_(\\w+)_(\\d+)","\\1",colnames(abund)),
                 Location=gsub("(\\w+)_(\\w+)_(\\d+)","\\2",colnames(abund)),
                 Date=gsub("(\\w+)_(\\w+)_(\\d+)","\\3",colnames(abund))))
write.table(stru,"../data/Sep2020_structure.txt",
            row.names = FALSE,quote=FALSE,col.names = FALSE,sep='\t')
div<-iDIP(abun = abund,struc=stru)
```

Let's reformat for improved interpretation

```{r iDipFormat}
div_dat<-data.frame(Within=c(NA,div["D_alpha.3",],div["D_alpha.2",],div["D_alpha.1",]),
           Between=c(NA,div["D_beta.3",],div["D_beta.2",],div["D_beta.1",]),
           Total=c(div["D_gamma",],div["D_gamma",],div["D_alpha.3",],div["D_alpha.2",]))
kable(div_dat)
```


### Hill numbers analysis


```{r calc_hill}
calc_hill<-function(vec,q){
  qD<-(sum(vec[vec>0]^q))^(1/(1-q))
  if(q==1){
    qD<-exp(-1*sum(vec[vec>0]*log(vec[vec>0])))
  }
  return(qD)
}
```

```{r reformatDates}
# only keep tows with > 0 observations
divdat<-combdat[rowSums(combdat[,6:ncol(combdat)])>0,]
divabunds<-as.matrix(divdat[rowSums(divdat[,6:ncol(divdat)])>0,6:ncol(divdat)])
# flip the dataset
divabunds<-t(divabunds)

# add column names
colnames(divabunds)<-divdat$Date


# combine samples from the same days 
# https://stackoverflow.com/questions/35626945/row-wise-sum-of-values-grouped-by-columns-with-same-name
abund<-t(rowsum(t(divabunds), group = colnames(divabunds), na.rm = T))
relabund<-sapply(1:ncol(abund),function(k)	abund[,k]/sum(abund[,k]))
colnames(relabund)<-colnames(abund)

```



```{r calcQs}
q0<-apply(relabund,2,calc_hill,q=0)
q1<-apply(relabund,2,calc_hill,q=1)
q2<-apply(relabund,2,calc_hill,q=2)

Qs<-data.frame(q0=q0,q1=q1,q2=q2)
Qs$date<-as.Date(gsub("(\\d+)_\\w+_\\w+","\\1",rownames(Qs)))

Qs<-Qs[order(Qs$date),]
```


```{r qsPlot_function}
# assumes the columns are q0,q1,q2 and lab_col is the name of the row with the x-axis labels
qsPlot<-function(Qs,lab_col,qcols=c("#a1dab4","#41b6c4","#225ea8")){

  plot(Qs$q0,type="l",xlab=lab_col,ylab="Species diversity",
     ylim=c(0,max(Qs$q0)+2),cex=2,lwd=2,axes=FALSE,col=qcols[1])
  axis(1,labels=c("",as.character(Qs[,lab_col])),at=0:nrow(Qs))
  axis(2,las=2,seq(-10,max(Qs$q0)+5,5))
  points(Qs$q0,pch=19,cex=2,lwd=2,col=qcols[1])
  text(1,Qs$q0[1]-1,"q=0",font = 3,col=qcols[1])
  
  points(Qs$q1,pch=16,col=qcols[2],cex=2,lwd=2)
  points(Qs$q1,lty=2,col=qcols[2],type='l',lwd=2)
  text(1,Qs$q1[1]+1,"q=1",col=qcols[2],font = 3)
  
  points(Qs$q2,pch=17,col=qcols[3],cex=2,lwd=2)
  points(Qs$q2,lty=3,col=qcols[3],type='l',lwd=2)
  text(1,Qs$q2[1]-1,"q=2",col=qcols[3],font = 3)
}
```

```{r plotDuvauchelleQs}

qsPlot(Qs,"date")

```


#### Compare seagrass vs bare 


```{r reformatWithMudvGrass}
# only keep tows with > 0 observations
divdat<-combdat[rowSums(combdat[,6:ncol(combdat)])>0,]
divabunds<-as.matrix(divdat[rowSums(divdat[,6:ncol(divdat)])>0,6:ncol(divdat)])
# flip the dataset
divabunds<-t(divabunds)

# add column names
colnames(divabunds)<-paste(divdat$Date,divdat$Season,divdat$SeagrassCover,sep="_")


# combine samples from the same days 
# https://stackoverflow.com/questions/35626945/row-wise-sum-of-values-grouped-by-columns-with-same-name
abund<-t(rowsum(t(divabunds), group = colnames(divabunds), na.rm = T))
relabund<-sapply(1:ncol(abund),function(k)	abund[,k]/sum(abund[,k]))
colnames(relabund)<-colnames(abund)

```


```{r calcQsMudvGrass,fig.width=7,fig.height=6,fig.keep='high',dpi=300}
q0<-apply(relabund,2,calc_hill,q=0)
q1<-apply(relabund,2,calc_hill,q=1)
q2<-apply(relabund,2,calc_hill,q=2)

Qs<-data.frame(q0=q0,q1=q1,q2=q2)
Qs$date<-as.Date(gsub("(\\d+)_\\w+_\\w+","\\1",rownames(Qs)))

Qs<-Qs[order(Qs$date),]
```


```{r DuvQs_MudvGrass,fig.width=8,fig.height=6,fig.keep='high',dpi=300}

par(mfrow=c(1,4),oma=c(2.5,2.5,2.5,1.5),mar=c(3,3,3,2))
qsPlot(Qs[grep("bare",rownames(Qs)),],"date")
legend("topleft","Bare",bty='n')

qsPlot(Qs[grep("sparse",rownames(Qs)),],"date")
legend("topleft","Sparse",bty='n')

qsPlot(Qs[grep("patchy",rownames(Qs)),],"date")
legend("topleft","Patchy",bty='n')

qsPlot(Qs[grep("dense",rownames(Qs)),],"date")
legend("topleft","Dense Seagrass",bty='n')

mtext("Species Diversity",2,outer=TRUE)
```


## Summaries

I can create a graph showing the numbers of species for each type of seagrass habitats.

```{r}
duvSummary<-data.frame(
  Status=paste(combdat$Season,combdat$seagrass_cover,sep="_"),
  Leptonotus=combdat$L.elevatus,
  Stigmatopora=combdat$S.nigra,
  Shrimp=combdat$Shrimp_spp,
  Flounder=combdat$Flounder,
  Crabs=rowSums(combdat[,25:30]),
  Snails=rowSums(combdat[,33:38]),
  OtherFish=rowSums(combdat[,17:23]),
  Other=rowSums(combdat[,c(31,32,39:34)])
  )

```

```{r}
duvCounts<-as.data.frame(duvSummary %>%
  group_by(Status) %>%
  summarise_all( sum))
```

```{r}

bareLept<-barplot(duvCounts[grep("Bare",duvCounts$Status),2],col = seasonCols)

barplot(duvCounts[grep("Bare",duvCounts$Status),3])


barplot(duvCounts[grep("Bare",duvCounts$Status),4],col = seasonCols)




```

```{r}
sppCols<-c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462')
spp<-colnames(duvSummary)[4:9]
```

```{r}
CountsPlot<-function(summaryMatrix, spp, sppCols,x.lim=c(0,7),y.lim=c(0,300)){
  
  
  plot(rep(1,nrow(summaryMatrix)),
       summaryMatrix[,1],
       xlim=x.lim,
       ylim=y.lim,
       xlab="",
       ylab="Counts",
       axes=F,
       type='n')
  axis(1,at=1:length(spp),labels=spp,lty = 0,las=2)
  axis(2,las=1)
  
  
  for(i in 1:length(spp)){
    points(rep(i,length(summaryMatrix[,spp[i]])),
       summaryMatrix[,spp[i]],
       pch=21,
       bg=scales::alpha(sppCols[i],0.75),
       col="dark grey",
       cex=2.5)
    lines(x=c(i-0.25,i+0.25),
          y=c(mean(summaryMatrix[,spp[i]]),
              mean(summaryMatrix[,spp[i]])),
          lwd=4,
          col=sppCols[i])
  }
}
```


```{r seagrassPlots}

par(mfrow=c(1,4))
CountsPlot(duvSummary[grep("Bare",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Sparse",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Patchy",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Dense",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)

```



```{r seasonPlots}

par(mfrow=c(1,4))
CountsPlot(duvSummary[grep("Spring",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Summer",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Autumn",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Winter",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)

```


```{r getImages}
bare <- image_read('../figs/seagrass_cover_images/Bare.PNG')
bare_png <- image_convert(bare, "png")

sparse <- image_read('../figs/seagrass_cover_images/Sparse.PNG')
sparse_png <- image_convert(sparse, "png")

patchy <- image_read('../figs/seagrass_cover_images/Patchy.PNG')
patchy_png <- image_convert(patchy, "png")

dense <- image_read('../figs/seagrass_cover_images/Dense.PNG')
dense_png <- image_convert(dense, "png")


image_append(image_scale(c(
  bare_png,
  sparse_png,
  patchy_png,
  dense_png
)))
```









