---
title: "Community analysis"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="../figs/")
```
```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(vegan)
library(indicspecies)
library(dplyr)
library(tidyr)
library(leaflet)
library(magick)
library(adespatial)
library(kableExtra)
```

```{r}
files.sources <- list.files(path = "../R/hierDiversity",full.names = TRUE)
tmp<-sapply(files.sources, source)

tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
```


```{r getData}
combdat<-read.csv("cleanedCombinedCommunity.csv",stringsAsFactors = TRUE)
combdat$SeagrassCover<-factor(combdat$SeagrassCover,
                              levels=c("bare","sparse","patchy","dense"))
combdat$Season<-factor(combdat$Season,
                       levels=c("Winter","Spring","Summer","Fall"))
```
```{r}
seasonCols<-c(Spring="coral",Summer="olivedrab3",Fall="gold",Winter="dodgerblue")
seagrassCs<-c(bare="#d9f0a3",sparse="#78c679",patchy="#41ab5d",dense="#005a32")
seasonPchs<-c(Spring=15,Summer=17,Fall=18,Winter=19)
seagrassPs<-c(bare=16, sparse=18, patchy=17, dense=15)
```

```{r getGPS}
gpsdat<-read.csv("../data/gpsdat.csv")
```

```{r}
lat <- c(-43.794,-43.752) #define our map's ylim
lon <- c(172.930,172.933) #define our map's xlim
center <- c(mean(c(gpsdat$lat_start,gpsdat$lat_stop)), 
            172.9336)  #tell what point to center on
zoom <- 16  #zoom: 1 = furthest out (entire globe), larger numbers = closer in

```



## Community analysis


```{r}
abunds<-as.matrix(combdat[rowSums(combdat[,6:ncol(combdat)])>4,
                          6:ncol(combdat)])
colnames(abunds)<-tolower(colnames(abunds))
ys<-combdat[rowSums(combdat[,6:ncol(combdat)])>4,
            c("SeagrassCover","Season","TowArea")]
densities<-abunds/combdat[rowSums(combdat[,6:ncol(combdat)])>4,"TowArea"]
# densities
reldensities<-decostand(densities,method="total",MARGIN = 1)
```


add space as a covariate using dbMEM -- this basically maxes it into a case of multiplying the sites/species matrix by the environment matrix and the matrix of locations. 


```{r spatialPackages}
library(ade4)
library(adespatial)
library(adegraphics)
library(spdep)
library(maptools)
```




First, I need to convert the lat/long coordinates to Cartesian X/Y coordinates
```{r convertXY}
# convert to XY coordinates to distances in meters
duvXY<-SoDA::geoXY(rowMeans(gpsdat[,grep("lat",colnames(gpsdat))]),
                   rowMeans(gpsdat[,grep("long",colnames(gpsdat))]), 
                   unit=1)
s.label(duvXY)
```

Then, I need to test for linear trends and detrend the data if the trend surface is significant. This means to regress all variables on the X – Y coordinates and retain the residuals. This can most easily be done using the function `lm()` (Borcard et al. 2011).

```{r}
duvXY<-duvXY[which(rowSums(combdat[,6:ncol(combdat)])>4),]
anova(rda(reldensities,duvXY))
# if significant, detrend
duvDetrend<-resid(lm(as.matrix(reldensities)~ ., data=as.data.frame(duvXY)))
```


### dbMEM

Use adespatial to identify the best threshold level

```{r, eval=FALSE}
thresh<-seq(50,100,10)

# convert neighborhoods to listw objects
listW<-listw.candidates(duvXY,nb='dnear',d2=thresh)

# apply the candidate neighborhood networks to the detrended data
duvThreshRes<-listw.select(x=duvDetrend,
                           candidates = listW)
saveRDS(duvThreshRes,"duvThreshRes.RDS")

```{r}
duvThreshRes$best.id
lwBest <- listW[[duvThreshRes$best.id]]
```

duvThreshRes<-readRDS("duvThreshRes.RDS")
```

Then use `dbmem` with that best threshold (`r duvThreshRes$best.id`).


```{r}
xydbmem<-as.data.frame(dbmem(duvXY,silent=FALSE,thresh=70))
# Create null model for forward selection, IF we have a significant result.
MEMrdaNULL <- capscale(duvDetrend ~ 1, data=xydbmem)
# Now run analysis on saturated dbMEM model
(MEMrda <- capscale(duvDetrend ~ ., data=xydbmem))

anova(MEMrda) # suggests spatial relationship 
```

Perform forward selection of the dbmem models. 

```{r}
forwardSpace <- ordiR2step(MEMrdaNULL, MEMrda, trace=FALSE)
forwardSpace$call
```

And combine the selected forward-selected axes with the space axes. 

```{r}
spaceAxes <- cbind(duvXY, xydbmem[, attr(forwardSpace$terms, "term.labels")])
```

Finally, use both environmental variables (separately) in the variance partitioning, along with space. 

```{r}
vp<-varpart(duvDetrend,ys$SeagrassCover,ys$Season,spaceAxes)
par(mfrow=c(1,2))
showvarparts(3,col=c("green", "red", "blue"))
plot(vp, digits=2, bg=c("green", "red", "blue"), Xnames = c("seagrass", "season","spatial"))
```

Because there are more spatial variables than either of the other explanatory variables, I need to only interpret the adjusted $R^2$ values. 

```{r}
vpTab<-dplyr::bind_rows(vp$part$fract,vp$part$indfract)
rownames(vpTab)<-c("seagrass",
                   "season",
                   "spatial",
                   "seagrass + season",
                   "seagrass + spatial",
                   "season + spatial",
                   "all",
                   "seagrass | season + spatial",
                   "season | seagrass + spatial",
                   "spatial | seagrass + season",
                   "intersection seagrass + season",
                   "intersection season + spatial",
                   "intersection seagrass + spatial",
                   "intersection seagrass + season + spatial",
                   "Residuals")
kable(vpTab,
      "latex",
      booktabs=TRUE,
      row.names = TRUE,
      caption="Results of variance partitioning using spatial variables as covariates.")
```

```{r saveVP}
write.csv(vpTab,"variationPartitioning.csv",row.names = TRUE,quote=FALSE)
```

Those that are marked as `testable=TRUE` [can be tested in an rda framework](https://www.davidzeleny.net/anadat-r/doku.php/en:varpart_examples), so we need rdas for the overall model, for each variable independently, and for each variable with conditioning effects of the others.

```{r}
global<-capscale(duvDetrend~.,data=bind_cols(spaceAxes,SeagrassCover=ys$SeagrassCover,Season=ys$Season))
space<-capscale(duvDetrend~.,data=spaceAxes)
seagrass<-capscale(duvDetrend~ys$SeagrassCover)
season<-capscale(duvDetrend~ys$Season)

spaceCond<-capscale(duvDetrend~. + Condition(ys$SeagrassCover) + Condition(ys$Season),data=spaceAxes)

# create a formula with the conditional space axes
form<-paste0("duvDetrend~ys$SeagrassCover + Condition(ys$Season)+",
              paste0("Condition(",colnames(spaceAxes),")",collapse = "+"))
seagrassCond<-capscale(formula(form),data=spaceAxes)
       
# now make seagrass conditional    
form<-paste0("duvDetrend~ys$Season + Condition(ys$SeagrassCover)+",
              paste0("Condition(",colnames(spaceAxes),")",collapse = "+"))              
seasonCond<-capscale(formula(form),data=spaceAxes)
anova(global)
anova(space)
anova(seagrass)
anova(season)
anova(spaceCond)
anova(seagrassCond)
anova(seasonCond)
```

All of these are significant at `P < 0.001`, suggesting that we can consider both the marginal and conditional components are relevant/significant.

```{r}
globalInt<-capscale(duvDetrend~as.matrix(spaceAxes)*ys$SeagrassCover*ys$Season)
globalSum<-summary(globalInt)
globalInt
```
```{r}
anova(globalInt)
```
scrs <- scores(globalInt, display = c("sites", "species"), scaling = 3)
sppScrs<-scrs$species
plotSpp<-as.data.frame(sppScrs[which(abs(sppScrs[,1])>=1 | abs(sppScrs[,2])>=1),])
```


```{r globalRDAplot}
par(mfrow=c(1,2),xpd=FALSE)
plot(globalInt,
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,2]*100,2),
       "% of constrained)"),
     xlim=c(-1.5,2),
     ylim=c(-1.5,1.5)
     )
points(globalInt,
       cex=1.5,
       col=scales::alpha(
         "dark gray",0.5
       ),
       bg=scales::alpha(
         "dark gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$Season,col=seasonCols[levels(ys$Season)],lwd=5)

arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.12,
       y1=plotSpp[,2]*0.12,
       length = 0.1,
       lwd=2)
text(x=plotSpp["shrimp",1]*0.11-0.1,
     y=plotSpp["shrimp",2]*0.11+0.2,
     labels = expression(italic("Caridea")))
text(x=plotSpp["flounder",1]*0.11+0.2,
     y=plotSpp["flounder",2]*0.11+0.2,
     expression(italic("Rhombosolea spp.")))
text(x=plotSpp["diloma",1]*0.14,
     y=plotSpp["diloma",2]*0.14,
     labels=expression(italic("D. subrostrata")))
text(x=plotSpp["mud_snail",1]*0.12+0.7,
     y=plotSpp["mud_snail",2]*0.12,
     labels = expression(italic("P. antipodarum")))
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",2]*0.11-0.1,
     labels = expression(italic("A. stutchburyi")))

legend("topright","A",cex=2,bty='n')
legend("bottom",c("Winter","Spring","Summer","Fall"),
       ncol=2,bty='n',
       col=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       bg=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       lwd=4,
       lty=1)

# 2nd plot

plot(globalInt,
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,2]*100,2),
       "% of constrained)"),
     xlim=c(-1.5,2),
     ylim=c(-2,2)
     )
points(globalInt,
       cex=1.5,
       col=scales::alpha(
         "dark gray",0.5
       ),
       bg=scales::alpha(
         "dark gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$SeagrassCover,col=seagrassCs[levels(ys$SeagrassCover)],lwd=5)
arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.12,
       y1=plotSpp[,2]*0.12,
       length = 0.1,
       lwd=2)
text(x=plotSpp["shrimp",1]*0.11-0.1,
     y=plotSpp["shrimp",2]*0.11+0.2,
     labels = expression(italic("Caridea")))
text(x=plotSpp["flounder",1]*0.11+0.2,
     y=plotSpp["flounder",2]*0.11+0.2,
     expression(italic("Rhombosolea spp.")))
text(x=plotSpp["diloma",1]*0.14,
     y=plotSpp["diloma",2]*0.14,
     labels=expression(italic("D. subrostrata")))
text(x=plotSpp["mud_snail",1]*0.12+0.7,
     y=plotSpp["mud_snail",2]*0.12,
     labels = expression(italic("P. antipodarum")))
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",2]*0.11-0.1,
     labels = expression(italic("A. stutchburyi")))

legend("topright","B",cex=2,bty='n')
legend("bottom",c("bare","sparse","patchy","dense"),
       ncol=2,
       lwd=4,
       lty=1,
       bty='n',
       col=seagrassCs[c("bare","sparse","patchy","dense")],
       bg=seagrassCs[c("bare","sparse","patchy","dense")])
```


```{r globalRDAplot2}
scrs <- scores(globalInt, display = "species", scaling = 3,choices = c(1,3,4))

par(mfrow=c(2,2),xpd=FALSE)
plotSpp<-as.data.frame(scrs[which(abs(scrs[,1])>=1 | abs(scrs[,2])>=1),])

plot(globalInt,
     choices=c(1,3),
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,3]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,3]*100,2),
       "% of constrained)"),
     )
points(globalInt,
       cex=1.5,
       choices=c(1,3),
       col=scales::alpha(
         "light gray",0.5
       ),
       bg=scales::alpha(
         "light gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$Season,
            col=seasonCols[levels(ys$Season)],
            lwd=5,choices=c(1,3))

arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.1,
       y1=plotSpp[,2]*0.1,
       length = 0.1,
       lwd=2)
text(x=plotSpp[c("shrimp","flounder"),1]*0.1+0.2,
     y=plotSpp[c("shrimp","flounder"),2]*0.1+0.2,
     labels = c("shrimp","flounder"))
text(x=plotSpp["diloma",1]*0.12,
     y=plotSpp["diloma",2]*0.12,
     labels="diloma")
text(x=plotSpp["mud_snail",1]*0.11+0.52,
     y=plotSpp["mud_snail",2]*0.11,
     labels = "mud_snail")
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",2]*0.11-0.1,
     labels = "austrovenus")


legend("topleft",c("Winter","Spring","Summer","Fall"),
       ncol=2,bty='n',
       col=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       bg=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       lwd=4,
       lty=1)

# 2nd plot

plot(globalInt,
     choices=c(1,3),
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,2]*100,2),
       "% of constrained)"),
     )
points(globalInt,
       cex=1.5,
       choices=c(1,3),
       col=scales::alpha(
         "light gray",0.5
       ),
       bg=scales::alpha(
         "light gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$SeagrassCover,
            col=seagrassCs[levels(ys$SeagrassCover)],
            lwd=5,choices=c(1,3))
arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.1,
       y1=plotSpp[,2]*0.1,
       length = 0.1,
       lwd=2)
text(x=plotSpp[c("shrimp","flounder"),1]*0.1+0.2,
     y=plotSpp[c("shrimp","flounder"),2]*0.1+0.2,
     labels = c("shrimp","flounder"))
text(x=plotSpp["diloma",1]*0.12,
     y=plotSpp["diloma",2]*0.12,
     labels="diloma")
text(x=plotSpp["mud_snail",1]*0.11+0.52,
     y=plotSpp["mud_snail",2]*0.11,
     labels = "mud_snail")
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",2]*0.11-0.1,
     labels = "austrovenus")


legend("topright",c("bare","sparse","patchy","dense"),
       ncol=2,
       lwd=4,
       lty=1,
       bty='n',
       col=seagrassCs[c("bare","sparse","patchy","dense")],
       bg=seagrassCs[c("bare","sparse","patchy","dense")])

# 4th axis
plotSpp<-as.data.frame(scrs[which(abs(scrs[,1])>=1 | abs(scrs[,3])>=1),])

plot(globalInt,
     choices=c(1,4),
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,4]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,4]*100,2),
       "% of constrained)"),
     )
points(globalInt,
       cex=1.5,
       choices=c(1,4),
       col=scales::alpha(
         "light gray",0.5
       ),
       bg=scales::alpha(
         "light gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$Season,
            col=seasonCols[levels(ys$Season)],
            lwd=5,choices=c(1,4))

arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.1,
       y1=plotSpp[,3]*0.1,
       length = 0.1,
       lwd=2)
text(x=plotSpp[c("shrimp","flounder"),1]*0.1+0.2,
     y=plotSpp[c("shrimp","flounder"),3]*0.1+0.2,
     labels = c("shrimp","flounder"))
text(x=plotSpp["diloma",1]*0.12,
     y=plotSpp["diloma",3]*0.12,
     labels="diloma")
text(x=plotSpp["mud_snail",1]*0.11+0.52,
     y=plotSpp["mud_snail",3]*0.11,
     labels = "mud_snail")
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",3]*0.11-0.1,
     labels = "austrovenus")


legend("topleft",c("Winter","Spring","Summer","Fall"),
       ncol=2,bty='n',
       col=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       bg=c(seasonCols[c("Winter","Spring","Summer","Fall")]),
       lwd=4,
       lty=1)

# 2nd plot

plot(globalInt,
     choices=c(1,4),
     display="sites",
     scaling=3,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(globalSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(globalSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(globalSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(globalSum$concont$importance[2,2]*100,2),
       "% of constrained)"),
     )
points(globalInt,
       cex=1.5,
       choices=c(1,4),
       col=scales::alpha(
         "light gray",0.5
       ),
       bg=scales::alpha(
         "light gray",0.5
       ),
       pch=19,
       display="sites",
       scaling=3)

ordiellipse(globalInt,ys$SeagrassCover,
            col=seagrassCs[levels(ys$SeagrassCover)],
            lwd=5,choices=c(1,4))
arrows(x0=rep(0,nrow(plotSpp)),
       y0=rep(0,nrow(plotSpp)),
       x1=plotSpp[,1]*0.1,
       y1=plotSpp[,3]*0.1,
       length = 0.1,
       lwd=2)
text(x=plotSpp[c("shrimp","flounder"),1]*0.1+0.2,
     y=plotSpp[c("shrimp","flounder"),3]*0.1+0.2,
     labels = c("shrimp","flounder"))
text(x=plotSpp["diloma",1]*0.12,
     y=plotSpp["diloma",3]*0.12,
     labels="diloma")
text(x=plotSpp["mud_snail",1]*0.11+0.52,
     y=plotSpp["mud_snail",3]*0.11,
     labels = "mud_snail")
text(x=plotSpp["austrovenus",1]*0.11+0.7,
     y=plotSpp["austrovenus",3]*0.11-0.1,
     labels = "austrovenus")


legend("topright",c("bare","sparse","patchy","dense"),
       ncol=2,
       lwd=4,
       lty=1,
       bty='n',
       col=seagrassCs[c("bare","sparse","patchy","dense")],
       bg=seagrassCs[c("bare","sparse","patchy","dense")])
```



### Analysis of multivariate patterns of association

```{r}
Patterns of abundance associated with seagrass cover:

coverIndics<-multipatt(abunds,ys$SeagrassCover)
summary(coverIndics)
```

```{r}
Patterns of abundance associated with season:

seasonIndics<-multipatt(abunds,ys$Season)
summary(seasonIndics)
```



## Diversity analysis


```{r reformatForDiversity}
# only keep tows with > 0 observations
divdat<-combdat[rowSums(combdat[,6:ncol(combdat)])>0,]
divabunds<-as.matrix(divdat[rowSums(divdat[,6:ncol(divdat)])>0,6:ncol(divdat)])
# relative
relabunds<-decostand(divabunds,method="total",MARGIN = 1)
# densities
divdensities<-divabunds/divdat$TowArea
```


### Rarefaction

```{r}
par(mfrow=c(1,2))
rarecurve(divabunds, main="By tow")
rarecurve(t(divabunds), 
          ylab="Number of individuals",
          main="By species")
```
```{r, eval=FALSE}
set.seed(8675309)
rars<-lapply(1:999,function(i,dat=divdat){
  dat<-divdat[,6:ncol(divdat)]
  ssz<-sum(dat[which.min(divdat$TowArea),])
  # rarefy
  rar<-suppressWarnings(rrarefy(dat,ssz))
  rownames(rar)<-paste(divdat$Date,
                       divdat$Season,
                       divdat$SeagrassCover,
                       sep="_")
  
  return(rar)
})

relrar<-lapply(rars,decostand,method="total",MARGIN=1)
saveRDS(relrar,"rarefiedCommunities.RDS")
```
```{r}
relrar<-readRDS("rarefiedCommunities.RDS")
```


### Hill numbers analysis

Within seagrass pairwise beta diversity


```{r}
# hierDiversity takes the non-relativized abundances
# least inclusive to most inclusive L->R
grps<-as.matrix(cbind(divdat[,c("SeagrassCover","Season")], 
            Site="Duvauchelle"))

divDensitiesQs<-lapply(list.files(pattern="divDensitiesQ.*RDS"),readRDS)

```

```{r getDivValues}
# the value is which row you want to extract, default is 1 (the mean)
getDivValues<-function(hierare){

  dat<-dplyr::bind_rows(lapply(hierare,function(hiers){
      estimates<-bind_rows(lapply(hiers,function(estimates){
        estimates$estimate<-rownames(estimates)
        estimates$level<-estimates$estimate[1]
        estimates<-pivot_longer(estimates,1:6,names_to = "diversity")
        return(estimates)
      }))
      return(estimates)
    }))

  return(dat)

}
```

```{r}
qdat<-lapply(divDensitiesQs,getDivValues)


for(q in 1:length(qdat)){
  qdat[[q]]$q<-q-1
}

qdat<-dplyr::bind_rows(qdat)
qdat$estimate[!qdat$estimate %in% c("SE","q0.025","q0.975")]<-"mu"
qdat$estimate[qdat$estimate=="SE"]<-"SEM"
actualQdat<-spread(qdat,estimate,value)

```


```{r, eval=FALSE}
rareQ0<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 0,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ0,"rarefied_Q0.RDS")

rareQ1<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 1,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ1,"rarefied_Q1.RDS")

rareQ2<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 2,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ2,"rarefied_Q2.RDS")
```


```{r}
extract_rarefied_estimates<-function(hdOutputRDS){
  rareQ<-readRDS(hdOutputRDS)
  estimatesQ<-dplyr::bind_rows(lapply(rareQ,function(hierare){
  
    dat<-dplyr::bind_rows(lapply(hierare,function(hiers){
      estimates<-bind_rows(lapply(hiers,function(estimates){
        return(estimates[1,])
      }))
      return(estimates)
    }))
    dat$level<-rownames(dat)
    
    longdat<-tidyr::pivot_longer(data=dat,colnames(dat)[1:6],
                                 names_to="diversity")
    return(longdat)
  }))
  
  meanQs<-tapply(estimatesQ$value,paste(estimatesQ$level,
                                        estimatesQ$diversity),
                 mean)
  meanQs<-data.frame(mu=meanQs,
                     level=gsub("(\\w+) (\\w+)","\\1",rownames(meanQs)),
                     diversity=gsub("(\\w+) (\\w+)","\\2",rownames(meanQs)),
                     row.names = NULL)
  meanQs$SEM<-tapply(estimatesQ$value,paste(estimatesQ$level,
                                            estimatesQ$diversity)
                     ,sem)
  meanQs$CIvalue<-tapply(estimatesQ$value,paste(estimatesQ$level,
                                                estimatesQ$diversity),
                         cim,output="value")
  return(meanQs)
}
```

```{r extractRarefiedEstimates}
rarefied_summaries<-lapply(list.files(pattern="rarefied_Q.*.RDS"),
                           extract_rarefied_estimates)
names(rarefied_summaries)<-list.files(pattern="rarefied_Q.*.RDS")

for(q in 1:length(rarefied_summaries)){
  rarefied_summaries[[q]]$q<-q-1
}

rarefied_summaries<-dplyr::bind_rows(rarefied_summaries)
```


```{r}
plotQs<-function(qdat,levs,cols,pchs,makeLeg=TRUE,ylims=NULL,SEM=TRUE){
  if(is.null(ylims)){
    ylims<-c(0,max(qdat$mu+0.2))
  }
  plot(0:2,seq(0,max(qdat$mu),length.out = 3),
       type="n",
     xlab="Order (q)",
     ylab="",
     xlim=c(-0.01,2.01),
     ylim=ylims,
     axes=FALSE,
     cex.lab=2)
  axis(1, pos=0,at=0:2,cex.axis=2)

  axis(2, las=1,cex.axis=2)
  for(grp in levs){
    subdat<-qdat[qdat$level %in% grp,]
    # background polygon
    if(isTRUE(SEM)){
      polygon(x = c(subdat$q,rev(subdat$q)),
            y=c(subdat$mu-subdat$SEM,
                rev(subdat$mu+subdat$SEM)),
            col=scales::alpha(cols[grp],0.5), 
            border=scales::alpha(cols[grp],0.5))
    } else{
      polygon(x = c(subdat$q,rev(subdat$q)),
            y=c(subdat$`q0.025`,
                rev(subdat$`q0.975`)),
            col=scales::alpha(cols[grp],0.5), 
            border=scales::alpha(cols[grp],0.5))
    }
    
    # add points
    points(subdat$q,subdat$mu,
           pch=pchs[grp],
           cex=2.5,
           col=cols[grp])
    
  }
  if(isTRUE(makeLeg)){
    legend("top",
             levs,
             col=cols[levs],
             pch=pchs[levs],
             ncol=length(levs)/2,
             bty='n',
             cex=2)
  }

}
```



```{r plotAlpha}
seasonsQ<-actualQdat[actualQdat$level %in% levels(combdat$Season) &
                               actualQdat$diversity == "alpha",]
seagrassQ<-actualQdat[actualQdat$level %in% levels(combdat$SeagrassCover) &
                               actualQdat$diversity == "alpha",]

par(mfrow=c(2,3),oma=c(1,1.5,3,0),mar=c(4,4,2.5,2),xpd=TRUE)
plotQs(seagrassQ,levels(combdat$SeagrassCover),
       seagrassCs,seagrassPs,FALSE,ylims=c(0,25))


mtext(expression("observed"~alpha),2,outer=FALSE,cex=1.5,line=3.5)
legend("topright","A",cex=2,bty='n')
legend(x=-0.25,y=33.5,xpd=TRUE,
       levels(combdat$SeagrassCover),
       col=seagrassCs[levels(combdat$SeagrassCover)],
       pch=seagrassPs[levels(combdat$SeagrassCover)],
       ncol=2,
       cex=2,
       bty='n')

plotQs(seasonsQ,levels(combdat$Season),seasonCols,
       seasonPchs,FALSE,ylims=c(0,25))
legend("topright","B",cex=2,bty='n')
legend(x=-0.2,y=33.5,xpd=TRUE,
       levels(combdat$Season),
       col=seasonCols[levels(combdat$Season)],
       pch=seasonPchs[levels(combdat$Season)],
       ncol=2,
       cex=2,
       bty='n')

plotQs(actualQdat[actualQdat$level %in% "Duvauchelle" &
                               actualQdat$diversity == "alpha",],
       "Duvauchelle",c(Duvauchelle="#36454f"),c(Duvauchelle=19),
       FALSE,ylims=c(0,25))
legend("topright","C",cex=2,bty='n')
legend(x=0,y=33.5,xpd=TRUE,
       "Overall",
       col="#36454f",
       pch=19,
       ncol=1,
       cex=2,
       bty='n')


# Rarefied ones
 seasonsQ<-rarefied_summaries[rarefied_summaries$level %in% levels(combdat$Season) &
                               rarefied_summaries$diversity == "alpha",]
seagrassQ<-rarefied_summaries[rarefied_summaries$level %in% levels(combdat$SeagrassCover) &
                               rarefied_summaries$diversity == "alpha",]

plotQs(seagrassQ,levels(combdat$SeagrassCover),
       seagrassCs,seagrassPs,FALSE,ylims=c(0,25))
legend("topright","D",cex=2,bty='n')
mtext(expression("rarefied"~alpha),2,outer=FALSE,cex=1.5,line=3.5)
plotQs(seasonsQ,levels(combdat$Season),
       seasonCols,seasonPchs,FALSE,ylims=c(0,25))
legend("topright","E",cex=2,bty='n')
plotQs(rarefied_summaries[rarefied_summaries$level %in% "Duvauchelle" &
                               rarefied_summaries$diversity == "alpha",],
       "Duvauchelle",c(Duvauchelle="#36454f"),c(Duvauchelle=19),
       FALSE,ylims=c(0,25))

legend("topright","F",cex=2,bty='n')
 


```


```{r plotPWturnover}

seasonsQ<-actualQdat[actualQdat$level %in% levels(combdat$Season) &
                               actualQdat$diversity == "pwTurnover",]
seagrassQ<-actualQdat[actualQdat$level %in% levels(combdat$SeagrassCover) &
                               actualQdat$diversity == "pwTurnover",]

par(mfrow=c(2,3),oma=c(1,4,3,0),mar=c(4,4,2.5,2),xpd=TRUE)
plotQs(seagrassQ,levels(combdat$SeagrassCover),
       seagrassCs,seagrassPs,FALSE,ylims=c(0,0.9))
legend("topright","A",cex=2,bty='n')
legend(x=-0.25,y=1.2,xpd=TRUE,
       levels(combdat$SeagrassCover),
       col=seagrassCs[levels(combdat$SeagrassCover)],
       pch=seagrassPs[levels(combdat$SeagrassCover)],
       ncol=2,
       cex=2,
       bty='n')

mtext("observed\npairwise turnover",2,outer=FALSE,cex=1.5,line=4)
plotQs(seasonsQ,levels(combdat$Season),
       seasonCols,seasonPchs,FALSE,ylims=c(0,0.9))
legend("topright","B",cex=2,bty='n')
legend(x=-0.25,y=1.2,xpd=TRUE,
       levels(combdat$Season),
       col=seasonCols[levels(combdat$Season)],
       pch=seasonPchs[levels(combdat$Season)],
       ncol=2,
       cex=2,
       bty='n')

plotQs(actualQdat[actualQdat$level %in% "Duvauchelle" &
                               actualQdat$diversity == "pwTurnover",],
       "Duvauchelle",c(Duvauchelle="#36454f"),c(Duvauchelle=19),
       FALSE,ylims=c(0,0.9))
legend("topright","C",cex=2,bty='n')
legend(x=0,y=1.2,xpd=TRUE,
       "Overall",
       col="#36454f",
       pch=19,
       ncol=1,
       cex=2,
       bty='n')

# Rarefied ones
seasonsQ<-rarefied_summaries[rarefied_summaries$level %in% levels(combdat$Season) &
                               rarefied_summaries$diversity == "pwTurnover",]
seagrassQ<-rarefied_summaries[rarefied_summaries$level %in% levels(combdat$SeagrassCover) &
                               rarefied_summaries$diversity == "pwTurnover",]

plotQs(seagrassQ,levels(combdat$SeagrassCover),
       seagrassCs,seagrassPs,FALSE,ylims=c(0,0.9))
legend("topright","D",cex=2,bty='n')
mtext("rarefied\npairwise turnover",2,outer=FALSE,cex=1.5,line=4)
plotQs(seasonsQ,levels(combdat$Season),
       seasonCols,seasonPchs,FALSE,ylims=c(0,0.9))
legend("topright","E",cex=2,bty='n')

plotQs(rarefied_summaries[rarefied_summaries$level %in% "Duvauchelle" &
                               rarefied_summaries$diversity == "pwTurnover",],
       "Duvauchelle",c(Duvauchelle="#36454f"),c(Duvauchelle=19),
       FALSE,ylims=c(0,0.9))
legend("topright","F",cex=2,bty='n')

```


