---
title: "Community analysis"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="../figs/")
```
```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(vegan)
library(indicspecies)
library(dplyr)
library(tidyr)
library(leaflet)
library(magick)
source("../R/iDIP.R")
```

```{r getData}
combdat<-read.csv("cleanedCombinedCommunity.csv",stringsAsFactors = TRUE)
levels(combdat$SeagrassCover)<-c("bare","sparse","patchy","dense")
```
```{r}
seasonCols<-c(Spring="coral",Summer="olivedrab3",Fall="gold",Winter="dodgerblue")
seagrassCs<-c(bare="coral",sparse="olivedrab3",patchy="gold",dense="dodgerblue")#c(bare="#ffffcc",sparse="#c2e699",patchy="#78c679",dense="#238443")
seasonPchs<-c(Spring=15,Summer=17,Fall=18,Winter=19)
seagrassPs<-c(bare=16, sparse=18, patchy=17, dense=15)
```


## Community analysis

(add space as a covariate)

### dbrda


```{r}
abunds<-as.matrix(combdat[rowSums(combdat[,6:ncol(combdat)])>4,6:ncol(combdat)])
ys<-combdat[rowSums(combdat[,6:ncol(combdat)])>4,c("SeagrassCover","Season","towArea")]
# change to area, but also maybe include nested date

seasonRDA<-capscale(abunds~SeagrassCover*Season + Condition(towArea), 
                 ys,
                 dist="bray")
```
```{r}
plot(seasonRDA)
anova(seasonRDA,by="terms")
seasonSum<-summary(seasonRDA)
```

These results suggest that season and seagrass cover -- and their interaction -- are important in determining the community in seagrass beds. 




```{r RDAplot, fig.cap="dbRDA plot of the first two axes, which describe >50% of the constrained variation. The colours of points represent seagrass cover categories (bare, sparse, patchy, dense) and the shapes of the points represent the season during which the sample was collected (winter, spring, summer, fall). Overlaid on top are the species, with flounder and shrimp being the most obvious species driving divergence."}

plot(seasonRDA,
     type='none',
     xlab=paste0(
       "CAP1 (",
       round(seasonSum$cont$importance[2,1]*100,2),
       "% of total; ", 
       round(seasonSum$concont$importance[2,1]*100,2),
       "% of constrained)"),
     ylab=paste0(
       "CAP2 (",
       round(seasonSum$cont$importance[2,2]*100,2),
       "% of total; ",
       round(seasonSum$concont$importance[2,2]*100,2),
       "% of constrained)")
     )
points(seasonRDA,
       cex=1.5,
       col=scales::alpha(
         seagrassCs[ys$SeagrassCover],0.5
       ),
       pch=seasonPchs[ys$Season])
text(seasonRDA, scaling = 2, display = "sp") 

par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0, 0, type='n', bty='n', xaxt='n', yaxt='n')
legend("topleft",c("Winter","Spring","Summer","Fall"),
       ncol=4,bty='n',
       pch=c(seasonPchs[c("Winter","Spring","Summer","Fall")]),
       col="grey")

legend("topright",c("bare","sparse","patchy","dense"),
       ncol=4,
       pch=19,
       bty='n',
       col=seagrassCs,
       bg=seagrassCs)
```

### Indicator species analysis

```{r}
coverIndics<-multipatt(abunds,ys$SeagrassCover)
summary(coverIndics)
```

```{r}
seasonIndics<-multipatt(abunds,ys$Season)
summary(seasonIndics)
```


#### exploring indicators

```{r getGPS}
gpsdat<-read.csv("../data/gpsdat.csv")
```

```{r}
lat <- c(-43.794,-43.752) #define our map's ylim
lon <- c(172.930,172.933) #define our map's xlim
center <- c(mean(c(gpsdat$lat_start,gpsdat$lat_stop)), 
            mean(c(gpsdat$long_start,gpsdat$long_stop)))  #tell what point to center on
zoom <- 16  #zoom: 1 = furthest out (entire globe), larger numbers = closer in

```


```{r makeSpeciesMap}
make_species_map<-function(gpsdat,speciesdat,speciesName,
                           center,zoom,color="yellow",
                           scale=5){
  long<-rowMeans(gpsdat[gpsdat$TowNum %in% speciesdat$TowNum,
                       c("long_start","long_stop")])
  lat<-rowMeans(gpsdat[gpsdat$TowNum %in% speciesdat$TowNum,
                       c("lat_start","lat_stop")])
  radius<-speciesdat[,speciesName]
  
  m <- leaflet() %>%
    addTiles() %>%  # Add default OpenStreetMap map tiles
    setView(lng=center[2], lat=center[1],zoom=zoom) %>% 
     #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
    addProviderTiles("Esri.WorldImagery") %>%
    addCircles(lng = long, lat = lat, weight = 1,
      radius = radius*scale,col=color)
  
  m
  return(m)
}

```

##### season indicators

```{r kahawai}
kahawai<-combdat[combdat$kahawai>0,]
make_species_map(gpsdat,kahawai,"kahawai",center,zoom)
```
```{r limpets}
limpets<-combdat[combdat$limpets>0,] # not abundant enough
make_species_map(gpsdat,limpets,"limpets",center,zoom)

```
```{r smelt}
# also patchy
smelt<-combdat[combdat$smelt>0,] # lots of smelt on one day
make_species_map(gpsdat,smelt,"smelt",center,zoom)

```

```{r bluecod}
blueCod<-combdat[combdat$blue_cod_juv>0,] # almost all from one day
make_species_map(gpsdat,blueCod,"blue_cod_juv",center,zoom)

```

```{r whitebait}
whitebait<-combdat[combdat$whitebait_below50mm>0,] # almost all from one day
make_species_map(gpsdat,whitebait,"whitebait_below50mm",center,zoom)

```

##### seagrass indicators

```{r arrowCrab}
# not bare
arrowCrab<-combdat[combdat$arrowheadCrab>0,]
make_species_map(gpsdat,arrowCrab,"arrowheadCrab",center,zoom)

```

```{r blenny}
blenny<-combdat[combdat$blenny>0,]
make_species_map(gpsdat,blenny,"blenny",center,zoom,scale=2)

```

```{r stigmatopora}
stigmatopora<-combdat[combdat$s.nigra>0,]
make_species_map(gpsdat,stigmatopora,"s.nigra",center,zoom)

```

```{r whelk}
mudWhelk<-combdat[combdat$mud_whelk>0,]
make_species_map(gpsdat,mudWhelk,"mud_whelk",center,zoom)

```

```{r spotty}
#sparse: also kahawai
spotty<-combdat[combdat$spotty>0,] 
make_species_map(gpsdat,spotty,"spotty",center,zoom)

```

```{r bristleworm}
bristleworm<-combdat[combdat$bristleworm>0,] 
make_species_map(gpsdat,bristleworm,"bristleworm",center,zoom)

```



## Diversity analysis


```{r reformatForDiversity}
# only keep tows with > 0 observations
divdat<-combdat[rowSums(combdat[,6:ncol(combdat)])>0,]
divabunds<-as.matrix(divdat[rowSums(divdat[,6:ncol(divdat)])>0,6:ncol(divdat)])
# relative
relabunds<-decostand(divabunds,method="total",MARGIN = 1)
# densities
divdensities<-divabunds/divdat$TowArea
```


### Rarefaction

```{r}
rarecurve(t(divabunds))
```

```{r, eval=FALSE}
set.seed(8675309)
rars<-lapply(1:999,function(i,dat=divdat){
  dat<-divdat[,6:ncol(divdat)]
  ssz<-sum(dat[which.min(divdat$TowArea),])
  # rarify
  rar<-suppressWarnings(rrarefy(dat,ssz))
  rownames(rar)<-paste(divdat$Date,
                       divdat$Season,
                       divdat$SeagrassCover,
                       sep="_")
  
  return(rar)
})

relrar<-lapply(rars,decostand,method="total",MARGIN=1)
saveRDS(relrar,"rarefiedCommunities.RDS")
```
```{r}
relrar<-readRDS("rarefiedCommunities.RDS")
```


### Hill numbers analysis

Within seagrass pairwise beta diversity
then you'd use those to calculate the upper levels?

Should my abundances be densities? - try it

```{r}
files.sources <- list.files(path = "../R/hierDiversity",full.names = TRUE)
tmp<-sapply(files.sources, source)

tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
```

```{r}
# hierDiversity takes the non-relativized abundances
grps<-as.matrix(cbind(divdat[,c("SeagrassCover","Season")], # least inclusive to most inclusive L->R
            Site="Duvauchelle"))
divActual<-hierDiversity(divabunds, 
              grps, 
              replace = NULL, 
              reps = 99, 
              q = 1,
              quant = c(0.025, 0.975),
              sims = FALSE)
```
saveRDS(divAcutal,"divActual.RDS")

divDensities<-hierDiversity(divdensities, 
              grps, 
              replace = NULL, 
              reps = 99, 
              q = 1,
              quant = c(0.025, 0.975),
              sims = FALSE)
saveRDS(divDensities,"divDensities.RDS")
```


```{r, eval=FALSE}
rareQ0<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 0,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ0,"rarefied_Q0.RDS")

rareQ1<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 1,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ1,"rarefied_Q1.RDS")

rareQ2<-lapply(rars,
               hierDiversity,
               group=grps,
               replace = NULL,
               reps = 1, 
               q = 2,
               quant = c(0.025, 0.975),
               sims = FALSE)
saveRDS(rareQ2,"rarefied_Q2.RDS")
```


```{r}
extract_rarefied_estimates<-function(hdOutputRDS){
  rareQ<-readRDS(hdOutputRDS)
  estimatesQ<-dplyr::bind_rows(lapply(rareQ,function(hierare){
  
    dat<-dplyr::bind_rows(lapply(hierare,function(hiers){
      estimates<-bind_rows(lapply(hiers,function(estimates){
        return(estimates[1,])
      }))
      return(estimates)
    }))
    dat$level<-rownames(dat)
    
    longdat<-tidyr::pivot_longer(data=dat,colnames(dat)[1:5],names_to="diversity")
    return(longdat)
  }))
  
  meanQs<-tapply(estimatesQ$value,paste(estimatesQ$level,estimatesQ$diversity),mean)
  meanQs<-data.frame(mu=meanQs,
                     level=gsub("(\\w+) (\\w+)","\\1",rownames(meanQs)),
                     diversity=gsub("(\\w+) (\\w+)","\\2",rownames(meanQs)),
                     row.names = NULL)
  meanQs$SEM<-tapply(estimatesQ$value,paste(estimatesQ$level,estimatesQ$diversity),sem)
  meanQs$CIvalue<-tapply(estimatesQ$value,paste(estimatesQ$level,estimatesQ$diversity),cim,output="value")
  return(meanQs)
}
```

```{r}
rarefied_summaries<-lapply(list.files(pattern="rarefied_Q.*.RDS"),extract_rarefied_estimates)
names(rarefied_summaries)<-list.files(pattern="rarefied_Q.*.RDS")
```



patchDiversity -- for a single level of groups (like seagrass)





```{r qsPlot_function}
# assumes the columns are q0,q1,q2 and lab_col is the name of the row with the x-axis labels
qsPlot<-function(Qs,lab_col,qcols=c("#a1dab4","#41b6c4","#225ea8")){

  plot(Qs$q0,type="l",xlab=lab_col,ylab="Species diversity",
     ylim=c(0,max(Qs$q0)+2),cex=2,lwd=2,axes=FALSE,col=qcols[1])
  axis(1,labels=c("",as.character(Qs[,lab_col])),at=0:nrow(Qs))
  axis(2,las=2,seq(-10,max(Qs$q0)+5,5))
  points(Qs$q0,pch=19,cex=2,lwd=2,col=qcols[1])
  text(1,Qs$q0[1]-1,"q=0",font = 3,col=qcols[1])
  
  points(Qs$q1,pch=16,col=qcols[2],cex=2,lwd=2)
  points(Qs$q1,lty=2,col=qcols[2],type='l',lwd=2)
  text(1,Qs$q1[1]+1,"q=1",col=qcols[2],font = 3)
  
  points(Qs$q2,pch=17,col=qcols[3],cex=2,lwd=2)
  points(Qs$q2,lty=3,col=qcols[3],type='l',lwd=2)
  text(1,Qs$q2[1]-1,"q=2",col=qcols[3],font = 3)
}
```




```{r DuvQs_Cover,fig.width=8,fig.height=6,fig.keep='high',dpi=300, eval=FALSE}

par(mfrow=c(1,4),oma=c(2.5,2.5,2.5,1.5),mar=c(3,3,3,2))
qsPlot(Qs[grep("bare",rownames(Qs)),],"date")
legend("topleft","Bare",bty='n')

qsPlot(Qs[grep("sparse",rownames(Qs)),],"date")
legend("topleft","Sparse",bty='n')

qsPlot(Qs[grep("patchy",rownames(Qs)),],"date")
legend("topleft","Patchy",bty='n')

qsPlot(Qs[grep("dense",rownames(Qs)),],"date")
legend("topleft","Dense Seagrass",bty='n')

mtext("Species Diversity",2,outer=TRUE)
```


## Summaries

I can create a graph showing the numbers of species for each type of seagrass habitats.

```{r}
duvSummary<-data.frame(
  Status=paste(combdat$Season,combdat$SeagrassCover,sep="_"),
  Leptonotus=combdat$leptonotus,
  Stigmatopora=combdat$s.nigra,
  Shrimp=combdat$shrimp,
  Flounder=combdat$flounder,
  Crabs=rowSums(combdat[,grep("crab",tolower(colnames(combdat)))]),
  Molluscs=rowSums(combdat[,c("Diloma","austrovenus","Lunella_smaragda","mud_whelk","whelk","paphies_australis")]),
  OtherFish=rowSums(combdat[,c("blenny","kahawai","spotty","wrasse","blue_cod_juv","whitebait_below50mm","smelt")]),
  Other=rowSums(combdat[,c("Seastar","sea_squirt","isopod","bristleworm")])
  )

```

```{r}
duvCounts<-as.data.frame(duvSummary %>%
  group_by(Status) %>%
  summarise_all( sum))
duvCounts$Season<-factor(gsub("(\\w+)_\\w+","\\1",duvCounts$Status),
                         levels=c("Winter","Spring","Summer","Fall"))
duvCounts<-duvCounts[,c("Status","Season","Leptonotus","Stigmatopora","Shrimp","Flounder","Crabs","Molluscs","OtherFish","Other")]
```



```{r}
# this plots each species over seasons
#requires a column called "Season" that is a factor. 
CountsPlot<-function(summaryMatrix, spp, sppCols,x.lim=c(0,7),y.lim=c(0,300)){
  
  # adjust the ylimit if necessary
  if(max(summaryMatrix[,spp])>max(y.lim)){
    y.lim[2]<-max(summaryMatrix[,spp])+50
  }
  # make an empty plot
  plot(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[1]],
       xlim=x.lim,
       ylim=y.lim,
       xlab="",
       ylab="Counts",
       axes=F,
       type='n',
       xpd=TRUE)
  axis(1,at=as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       labels=levels(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       lty = 0,las=2,cex=1)
  axis(2,las=1,cex=1)
  
  
  for(i in 1:length(spp)){
    # add points for each species
    points(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[i]],
       col=sppCols[i],
       lwd=2,
       type = "l")
    points(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[i]],
       pch=21,
       bg=scales::alpha(sppCols[i],0.75),
       col="dark grey",
       cex=2.5)
  }
}
```


```{r seagrassPlots}
spp<-colnames(duvCounts)[3:ncol(duvCounts)]
sppCols<-c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5')
names(sppCols)<-duvCounts[,3:ncol(duvCounts)]

par(mfrow=c(1,4),mar=c(5,3,2,0),xpd=TRUE,oma=c(0,0,0,0))
CountsPlot(duvCounts[grep("bare",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("sparse",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("patchy",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("dense",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
```

```{r}
# plot the relative abundances

duvRelSum<-data.frame(
  Status=paste(divdat$Season,divdat$SeagrassCover,sep="_"),
  Leptonotus=relabunds[,"leptonotus"],
  Stigmatopora=relabunds[,"s.nigra"],
  Shrimp=relabunds[,"shrimp"],
  Flounder=relabunds[,"flounder"],
  Crabs=rowSums(relabunds[,grep("crab",tolower(colnames(relabunds)))]),
  Molluscs=rowSums(relabunds[,c("Diloma","austrovenus","Lunella_smaragda","mud_whelk","whelk","paphies_australis")]),
  OtherFish=rowSums(relabunds[,c("blenny","kahawai","spotty","wrasse","blue_cod_juv","whitebait_below50mm","smelt")]),
  Other=rowSums(relabunds[,c("Seastar","sea_squirt","isopod","bristleworm")])
  )

```

```{r seasonPlots}
```{r}
duvCounts<-as.data.frame(duvRelSum %>%
  group_by(Status) %>%
  summarise_all( sum))
duvCounts$Season<-factor(gsub("(\\w+)_\\w+","\\1",duvCounts$Status),
                         levels=c("Winter","Spring","Summer","Fall"))
duvCounts<-duvCounts[,c("Status","Season","Leptonotus","Stigmatopora","Shrimp","Flounder","Crabs","Molluscs","OtherFish","Other")]
```

par(mfrow=c(1,4))
CountsPlot(duvSummary[grep("Spring",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Summer",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Autumn",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)
CountsPlot(duvSummary[grep("Winter",duvSummary$Status),4:9],spp=spp,sppCols = sppCols)

```


```{r getImages}
bare <- image_read('../figs/seagrass_cover_images/Bare.PNG')
bare_png <- image_convert(bare, "png")

sparse <- image_read('../figs/seagrass_cover_images/Sparse.PNG')
sparse_png <- image_convert(sparse, "png")

patchy <- image_read('../figs/seagrass_cover_images/Patchy.PNG')
patchy_png <- image_convert(patchy, "png")

dense <- image_read('../figs/seagrass_cover_images/Dense.PNG')
dense_png <- image_convert(dense, "png")


image_append(image_scale(c(
  bare_png,
  sparse_png,
  patchy_png,
  dense_png
)))
```









