---
title: "Summarizing sampling effort and data"
author: "Sarah Flanagan"
date: "23/08/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
browsable(
   tagList(
      list(
         tags$head(
            tags$style(
               ".leaflet .legend {
                 line-height: 30px;
                 font-size: 30px;
                 }",
              ".leaflet .legend i{
                width: 200px;
                height: 200px;
                 }"
            )
         ),
       leaf)))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="../figs/")
```
```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(vegan)
library(dplyr)
library(tidyr)
library(leaflet)
library(magick)
library(adespatial)
library(kableExtra)
```

```{r}
files.sources <- list.files(path = "../R/hierDiversity",full.names = TRUE)
tmp<-sapply(files.sources, source)

tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
```


```{r getData}
combdat<-read.csv("cleanedCombinedCommunity.csv",stringsAsFactors = TRUE)
combdat$SeagrassCover<-factor(combdat$SeagrassCover,
                              levels=c("bare","sparse","patchy","dense"))
combdat$Season<-factor(combdat$Season,
                       levels=c("Winter","Spring","Summer","Fall"))
```
```{r colsPchs}
seasonCols<-c(Spring="coral",Summer="olivedrab3",Fall="gold",Winter="dodgerblue")
seagrassCs<-c(bare="#d9f0a3",sparse="#78c679",patchy="#41ab5d",dense="#005a32")
seasonPchs<-c(Spring=15,Summer=17,Fall=18,Winter=19)
seagrassPs<-c(bare=16, sparse=18, patchy=17, dense=15)
```
```{r sppOrder}
sppOrder<-c(
            "flounder",
            "s.nigra",
            "leptonotus",
            "blue_cod_juv",
            "kahawai",
            "whitebait_below50mm",
            "smelt",
            "spotty",
            "wrasse",
            "blenny",
            "paphies_australis",
            "austrovenus",
            "shrimp",
            "swimmer_crab",
            "Pie_crust_crab",
            "hermit_crab",
            "mud_crab",
            "stalkeyed_crab",
            "arrowheadCrab",
            "Lunella_smaragda",
            "mud_whelk",
            "whelk",
            "mud_snail",
            "Diloma",
            "Seastar",
            "limpet",
            "bristleworm",
            "sea_squirt",
            "isopod"
            )
```


```{r getGPS}
gpsdat<-read.csv("../data/gpsdat.csv")
```

```{r centreGPS}

center <- c(mean(c(gpsdat$lat_start,gpsdat$lat_stop)), 
           172.9336)  #tell what point to center on
zoom <- 17  #zoom: 1 = furthest out (entire globe), larger numbers = closer in

```

## Sampling distribution and effort

We sampled on `r length(levels(combdat$Date))` occasions throughout the four seasons from October 2019 through October 2020. Each date had at least `r min(tapply(combdat$TowNum,combdat$Date,length))` individual tows through seagrass beds, with a total area covered across all sampling of `r sum(combdat$TowArea)`. 

```{r towSum}
TowSummary<-bind_cols(pivot_wider(combdat, 
                        id_cols = 1:2,
                        names_from = "SeagrassCover",
                        values_from = "TowArea",
                        names_prefix="N_",
                        values_fn = list(TowArea = length)),
                      pivot_wider(combdat, 
                        id_cols = 1:2,
                        names_prefix="Mean_",
                        names_from = "SeagrassCover",
                        values_from = "TowArea",
                        values_fn = list(TowArea = mean)),
                      pivot_wider(combdat, 
                        id_cols = 1:2,
                        names_prefix="SEM_",
                        names_from = "SeagrassCover",
                        values_from = "TowArea",
                        values_fn = list(TowArea = sem))
                      )
TowSummary<-TowSummary[,-c(7,8,13,14)]
colnames(TowSummary)[1:2]<-c("Date","Season")
# rearrange column order
TowSummary<-TowSummary[,c(1,2,
                          grep("bare",colnames(TowSummary)),
                          grep("sparse",colnames(TowSummary)),
                          grep("patchy",colnames(TowSummary)),
                          grep("dense",colnames(TowSummary)))]

kable(TowSummary,"latex",booktab=TRUE,
      caption="Table of the number of tows (N), mean tow area, and standard error of the mean (SEM) tow area for each seagrass cover type on each sampling date.")

```
```{r saveTowSummary,eval=FALSE}
write.csv(TowSummary,"TowSummary.csv",row.names = FALSE,quote=FALSE)
```

We can also plot the locations of the tows on the map.

```{r mapCover, fig.height=2,fig.width=4.5,res=300}
  
m<- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(lng=center[2], lat=center[1],zoom=zoom) %>% 
  
  #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
  addProviderTiles("Esri.WorldImagery") %>%
  addLegend(position="topright",
                  colors = seagrassCs, 
                  labels = names(seagrassCs),
                  title = "Seagrass cover",
                  opacity = 1
  ) %>%  
  addMiniMap(tiles="Esri.WorldImagery",
             width=200, height=200, position="bottomright") 
for(i in 1:nrow(gpsdat)){
  color<-seagrassCs[[combdat$SeagrassCover[combdat$TowNum==gpsdat[i,"TowNum"]]]]
  m <- addPolylines(m, lat = as.numeric(gpsdat[i, c("lat_start", "lat_stop")]),
                    lng = as.numeric(gpsdat[i, c("long_start", "long_stop")]),
                    color = color)
}

  m
```


```{r mapSeason, fig.height=2,fig.width=4.5,res=300}
  
seasonCs<-seasonCols
seasonCs[["Summer"]]<-"#9acd32"
m<- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(lng=center[2], lat=center[1],zoom=zoom) %>% 
  
  #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
  addProviderTiles("Esri.WorldImagery") %>%
  addLegend("topright",colors = seasonCs, labels = names(seasonCs),
    title = "Season",
    opacity = 1
  ) %>%  
  addMiniMap(tiles="Esri.WorldImagery",
             height=200,width=200, position="bottomright")
for(i in 1:nrow(gpsdat)){
  color<-seasonCs[[combdat$Season[combdat$TowNum==gpsdat[i,"TowNum"]]]]
  m <- addPolylines(m, lat = as.numeric(gpsdat[i, c("lat_start", "lat_stop")]),
                    lng = as.numeric(gpsdat[i, c("long_start", "long_stop")]),
                    color = color)
}

  m
```

We'll also create a map of NZ with Duvauchelle highlighted

```{r largeMap}
m<- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(lng=center[2], lat=center[1],zoom=5) %>% 
  
  #addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
  addProviderTiles("Esri.WorldImagery") %>%  
  addRectangles(lng1 = 172.9,#min(c(gpsdat$long_start,gpsdat$long_stop)),
  lat1 = -43.7,#min(c(gpsdat$lat_start,gpsdat$lat_stop)),
  lng2 = 173,#max(c(gpsdat$long_start,gpsdat$long_stop)),
  lat2 = -43.8,#max(c(gpsdat$lat_start,gpsdat$lat_stop)),
  color = "red") %>%
  addMiniMap(zoomLevelFixed = 12,tiles="Esri.WorldImagery",
             width=200, height=200, position="bottomright") 


  m
```

I'll then stitch the static images together in GIMP.

## Overall counts data summary

```{r makeOutputCounts}
outputCounts<- as.data.frame(dplyr::bind_rows(
  lapply(colnames(combdat[,-c(1:5)]),function(column){
    tapply(combdat[,column], combdat$Season,sum)
  }))
)
rownames(outputCounts)<-colnames(combdat[,-c(1:5)])
write.csv(outputCounts,"TotalCountsAll.csv",quote=FALSE)
```

```{r outputCountsByBoth}
outputCountsSxS<- as.data.frame(dplyr::bind_rows(
  lapply(colnames(combdat[,-c(1:5)]),function(column){
    tapply(combdat[,column], paste0(combdat$Season,"_",combdat$SeagrassCover),sum)
  }))
)
rownames(outputCountsSxS)<-colnames(combdat[,-c(1:5)])
colnames(outputCountsSxS)<-paste0(colnames(outputCountsSxS),"_N")
outputCountsSxS<-outputCountsSxS[sppOrder,]

densities<-combdat[,6:ncol(combdat)]/combdat[,"TowArea"]
densityMeans<-as.data.frame(dplyr::bind_rows(
  lapply(colnames(densities),function(column){
    tapply(densities[,column], paste0(combdat$Season,"_",combdat$SeagrassCover),mean)
  }))
)
rownames(densityMeans)<-colnames(densities)
colnames(densityMeans)<-paste0(colnames(densityMeans),"_meanDensity")
densityMeans<-densityMeans[sppOrder,]

densitySEMs<-as.data.frame(dplyr::bind_rows(
  lapply(colnames(densities),function(column){
    tapply(densities[,column], paste0(combdat$Season,"_",combdat$SeagrassCover),sem)
  }))
)
rownames(densitySEMs)<-colnames(densities)
colnames(densitySEMs)<-paste0(colnames(densitySEMs),"_semDensity")
densitySEMs<-densitySEMs[sppOrder,]

SxSsummary<-as.data.frame(dplyr::bind_cols(outputCountsSxS,densityMeans,densitySEMs))


# reorder columns
levCombos<-NULL
for(season in levels(combdat$Season)){
  for(cover in levels(combdat$SeagrassCover)){
    levCombos<-c(levCombos,paste0(season,"_",cover))
  }
}
colOrder<-unlist(lapply(levCombos,grep,x=colnames(SxSsummary)))
SxSsummary<-SxSsummary[,colOrder]


```
```{r saveSummaries}
write.csv(SxSsummary[,grep("N",colnames(SxSsummary))],
          "SeagrassSeasonCounts.csv",quote=FALSE,row.names = TRUE)
write.csv(SxSsummary[,-grep("N",colnames(SxSsummary))],
          "SeagrassSeasonDensities.csv",quote=FALSE,row.names = TRUE)
write.csv(SxSsummary,"SeagrassSeasonSummary.csv",quote=FALSE,row.names = TRUE)
```


## High-level category summaries

I can create a graph showing the numbers of species for each type of seagrass habitats.

```{r duvSummary}
duvSummary<-data.frame(
  Status=paste(combdat$Season,combdat$SeagrassCover,sep="_"),
  Leptonotus=combdat$leptonotus,
  Stigmatopora=combdat$s.nigra,
  Shrimp=combdat$shrimp,
  Flounder=combdat$flounder,
  Crabs=rowSums(combdat[,grep("crab",tolower(colnames(combdat)))]),
  Molluscs=rowSums(combdat[,c("Diloma","austrovenus","Lunella_smaragda","mud_whelk","whelk","paphies_australis")]),
  OtherFish=rowSums(combdat[,c("blenny","kahawai","spotty","wrasse","blue_cod_juv","whitebait_below50mm","smelt")]),
  Other=rowSums(combdat[,c("Seastar","sea_squirt","isopod","bristleworm")])
  )

```

```{r duvCounts}
duvCounts<-as.data.frame(duvSummary %>%
  group_by(Status) %>%
  summarise_all( sum))
duvCounts$Season<-factor(gsub("(\\w+)_\\w+","\\1",duvCounts$Status),
                         levels=c("Winter","Spring","Summer","Fall"))
duvCounts<-duvCounts[,c("Status","Season","Leptonotus","Stigmatopora","Shrimp","Flounder","Crabs","Molluscs","OtherFish","Other")]
```

```{r saveCountsSummary}
outputCounts<- as.data.frame(dplyr::bind_rows(
  lapply(colnames(duvCounts[,-c(1:2)]),function(column){
    tapply(duvCounts[,column], duvCounts$Season,sum)
  }))
)
rownames(outputCounts)<-colnames(duvCounts[,-c(1:2)])
write.csv(outputCounts,"TotalCountsSummary.csv",quote=FALSE)
```



```{r CountsPlotFXN}
# this plots each species over seasons
#requires a column called "Season" that is a factor. 
CountsPlot<-function(summaryMatrix, spp, sppCols,x.lim=c(0,7),y.lim=c(0,300)){
  
  # adjust the ylimit if necessary
  if(max(summaryMatrix[,spp])>max(y.lim)){
    y.lim[2]<-max(summaryMatrix[,spp])+50
  }
  # make an empty plot
  plot(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[1]],
       xlim=x.lim,
       ylim=y.lim,
       xlab="",
       ylab="",
       cex.lab=2,
       axes=F,
       type='n',
       xpd=TRUE)
  axis(1,at=as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       labels=levels(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       lty = 0,las=2,cex.axis=2)
  axis(2,las=1,cex.axis=2)
  
  
  for(i in 1:length(spp)){
    # add points for each species
    points(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[i]],
       col=sppCols[i],
       lwd=2,
       type = "l")
    points(as.numeric(summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),"Season"]),
       summaryMatrix[order(as.numeric(summaryMatrix[,"Season"])),spp[i]],
       pch=21,
       bg=scales::alpha(sppCols[i],0.75),
       col="dark grey",
       cex=3)
  }
}
```


```{r seagrassPlotSettings}
spp<-colnames(duvCounts)[3:ncol(duvCounts)]
sppCols<-c('#8dd3c7','#ffffb3','#b3de69','#80b1d3','#fb8072','#fdb462','#bebada','#fccde5')
names(sppCols)<-duvCounts[,3:ncol(duvCounts)]

```


```{r seagrassPlots}
par(mfrow=c(1,4),mar=c(5,3,2,0),xpd=TRUE,oma=c(0,0,0,0))
CountsPlot(duvCounts[grep("bare",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("sparse",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("patchy",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
CountsPlot(duvCounts[grep("dense",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,3500))
```


```{r reformatForDiversity}
# only keep tows with > 0 observations
divdat<-combdat[rowSums(combdat[,6:ncol(combdat)])>0,]
divabunds<-as.matrix(divdat[rowSums(divdat[,6:ncol(divdat)])>0,6:ncol(divdat)])
# relative
relabunds<-decostand(divabunds,method="total",MARGIN = 1)
# densities
divdensities<-divabunds/divdat$TowArea
```

```{r relativeSummary}
# plot the relative abundances

duvRelSum<-data.frame(
  Status=paste(divdat$Season,divdat$SeagrassCover,sep="_"),
  Leptonotus=relabunds[,"leptonotus"],
  Stigmatopora=relabunds[,"s.nigra"],
  Shrimp=relabunds[,"shrimp"],
  Flounder=relabunds[,"flounder"],
  Crabs=rowSums(relabunds[,grep("crab",tolower(colnames(relabunds)))]),
  Molluscs=rowSums(relabunds[,c("Diloma","austrovenus","Lunella_smaragda","mud_whelk","whelk","paphies_australis")]),
  OtherFish=rowSums(relabunds[,c("blenny","kahawai","spotty","wrasse","blue_cod_juv","whitebait_below50mm","smelt")]),
  Other=rowSums(relabunds[,c("Seastar","sea_squirt","isopod","bristleworm")])
  )

```

```{r duvCountsOrdered}
duvCounts<-as.data.frame(duvRelSum %>%
  group_by(Status) %>%
  summarise_all( sum))
duvCounts$Season<-factor(gsub("(\\w+)_\\w+","\\1",duvCounts$Status),
                         levels=c("Winter","Spring","Summer","Fall"))
duvCounts<-duvCounts[,c("Status","Season","Leptonotus","Stigmatopora","Shrimp","Flounder","Crabs","Molluscs","OtherFish","Other")]
```




```{r getImages}
bare <- image_read('../figs/seagrass_cover_images/Bare.PNG')
bare_png <- image_convert(bare, "png")

sparse <- image_read('../figs/seagrass_cover_images/Sparse.PNG')
sparse_png <- image_convert(sparse, "png")

patchy <- image_read('../figs/seagrass_cover_images/Patchy.PNG')
patchy_png <- image_convert(patchy, "png")

dense <- image_read('../figs/seagrass_cover_images/Dense.PNG')
dense_png <- image_convert(dense, "png")

seagrassPhotos<-image_append(image_scale(c(bare_png,sparse_png,patchy_png,dense_png)))
```




```{r relativeAbundanceFig}
png("../figs/relativeAbundanceSummaries.png",width = 1512, height = 400, res = 96)
par(mfrow=c(1,4), oma=c(4,0,2,0), mar=c(4,5.5,1,0),xpd=TRUE)
CountsPlot(duvCounts[grep("bare",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,30))
#rasterImage(bare_png,xleft = 0,xright = 5,ytop = -26,ybottom=-30,xpd=TRUE)
mtext("Relative abundance",side=2,outer=FALSE,line=3,cex=1.5)

CountsPlot(duvCounts[grep("sparse",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,30))
#rasterImage(sparse_png,xleft = 0,xright = 5,ytop = -26,ybottom=-30,xpd=TRUE)

CountsPlot(duvCounts[grep("patchy",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,30))
#rasterImage(patchy_png,xleft = 0,xright = 5,ytop = -26,ybottom=-30,xpd=TRUE)

CountsPlot(duvCounts[grep("dense",duvCounts$Status),],spp=spp,sppCols = sppCols,y.lim = c(0,30))
#rasterImage(dense_png,xleft = 0,xright = 5,ytop = -26,ybottom=-30,xpd=TRUE)


outer_legend("top",spp,col=sppCols,pch=21,pt.bg=sppCols,lty=1,bty='n',ncol=8,cex=2,lwd=2,pt.cex=2)
dev.off()
```


```{r saveRelAbundFig}
relfig<-image_convert(image_read('../figs/relativeAbundanceSummaries.png'),"png")
combinedSummary<-image_append(c(relfig,seagrassPhotos), stack = TRUE)
image_write(combinedSummary,
            path = "../figs/relativeAbundanceSeagrass.png",
            format = "png")
```


