---
title: "Data processing"
author: "Sarah Flanagan"
date: "16/06/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "./figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="./figs/")
```

```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(dplyr)
library(tidyr)
library(geosphere)
```

# Community data

```{r}
rawdat<-read.csv("../data/MonthlyRawSampling_Duvauchelle_2020.csv")

# remove unnecessary column
rawdat<-subset(rawdat,select=-X)

```

First, check for NAs that should be zeros

```{r}
rawdat[,17:62][is.na(rawdat[,17:62])]<-0
```

Then add a date column

```{r}
rawdat$Day<-as.Date(
  paste0(
    gsub("^(\\d)$","0\\1",rawdat$Date),
    gsub("^(\\d)$","0\\1",rawdat$Month),
    rawdat$Year),
  format = "%d%m%Y")
```

And group them by season

```{r}
rawdat$Season<-gsub("\\d+-(\\d+)-\\d+","\\1",rawdat$Day)
rawdat$Season[rawdat$Season %in% c("09","10","11")]<-"Spring"
rawdat$Season[rawdat$Season %in% c("12","01","02")]<-"Summer"
rawdat$Season[rawdat$Season %in% c("03","04","05")]<-"Fall"
rawdat$Season[rawdat$Season %in% c("06","07","08")]<-"Winter"
```


Calculate the area of the tows

```{r}
towArea<-apply(rawdat,1,function(dat){
  towArea<-distm(c(as.numeric(dat["long_start"]),
          as.numeric(dat["lat_start"])),
        c(as.numeric(dat["long_stop"]),
          as.numeric(dat["lat_stop"])))
  towArea<-towArea*1.2
  return(towArea)
})
# turn zeroes into 1s
towArea[which(towArea==0)]<-1

# add to data sheet
rawdat$towArea<-towArea
```

Finally remove any rows with missing seagrass covers

```{r}
rawdat<-rawdat[!is.na(rawdat$seagrass_cover),]
```

Fix inconsistent seagrass cover typing

```{r}
rawdat$seagrass_cover<-tolower(rawdat$seagrass_cover)
rawdat$seagrass_cover[rawdat$seagrass_cover=="spare"]<-"sparse"
```

Remove commas from Comments column

```{r}
rawdat$Comments<-gsub(",","",rawdat$Comments)
```

```{r}
rawdat$TowNum<-rownames(rawdat)
```


```{r}
write.csv(rawdat,"../data/cleanedCommunity.csv",quote=FALSE, row.names = FALSE)
```


Combine columns so that each organism has its a single column

```{r}
combNames<-grep("mm$",colnames(rawdat),value=TRUE)
combNames<-combNames[combNames!="whitebait_below50mm"] # only whitebait instance
combPatts<-unique(gsub("_.*$","",combNames))
combPatts[combPatts=="austrovenus"]<-"venus" # one included 'venus_bivalve'

totals<-as.data.frame(do.call(cbind,lapply(combPatts,function(pattern,dat){
  tot<-rowSums(dat[,grep(pattern,colnames(dat))])
  return(tot)
},dat=rawdat)))
colnames(totals)<-combPatts
colnames(totals)[colnames(totals)=="venus"]<-"austrovenus"
totals$isopod<-rowSums(rawdat[,c("isopod","green_centipede")])

nonCombNames<-colnames(rawdat[,17:62])[!colnames(rawdat[,17:62]) %in% combNames]
nonCombNames<-grep("wrasse",nonCombNames,invert=TRUE,value=TRUE)
nonCombNames<-grep("venus",nonCombNames,invert=TRUE,value=TRUE)
nonCombNames<-nonCombNames[!nonCombNames %in% c("isopod","green_centipede")]
```



Create summary data frame

```{r}
combdat<-data.frame(cbind(
  Date=rawdat$Day,
  Season=rawdat$Season,
  TowNum=rawdat$TowNum,
  TowArea=rawdat$towArea,
  SeagrassCover=rawdat$seagrass_cover,
  totals,
  rawdat[,nonCombNames]
  )
)
```

```{r, eval=FALSE}
write.csv(combdat,"cleanedCombinedCommunity.csv",quote=FALSE, row.names = FALSE)
```
```{r}
combdat<-read.csv("cleanedCombinedCommunity.csv")
```

```{r}
gpsdat<-cbind(TowNum=rownames(rawdat),
              rawdat[,c("long_start","long_stop","lat_start","lat_stop")])
write.csv(gpsdat,"../data/gpsdat.csv",quote=FALSE,row.names=FALSE)
```


## Group to common taxonomic level

Some animals we identified to species, but others we grouped at higher orders. Shrimp in particular were a challenge -- the shrimp species likely included in our counts include:

- *Palaemon affinis* (glass shrimp), family Palaemonidae, infraorder Caridea
- *Alphaeus novazealandiae* or *A. socialis* (snapping shrimp | kowhitiwhiti moana), family alpheidae, infraorder caridea
- *Alope spinifrons* (painted shrimp), family Hippolytidae, infraorder Caridea
- *Pontophilus* or *Parapontophilus* species, infraorder Caridea
- *Hippolyte* spp. (chamaeleon shrimp), family Hippolytidae, infraorder Caridea

Which means infraorder is the lowest taxonomic unit for these animals. To summarize all of our species, see \@ref(tab:taxonomicTable).

```{r taxonomicTable}
names<-read.csv("../data/species_names.csv")
kable(names,booktabs=TRUE)
```


To ensure that the varying levels of lowest taxonomic identification don't impact the results, I'll create a data.frame where all of the animals have been grouped at the level of Order -- so I'll sum up any of the obervations at lower levels.  

```{r}
orders<-combdat
names<-names[match(colnames(orders[,4:ncol(orders)]),names$ID ),]
colnames(orders)[4:ncol(orders)]<-as.character(names$order)

ordAbund<-do.call(cbind,lapply( levels(as.factor(names$order)), function(name,dat){
  
  if(length(which((colnames(dat) %in% name)==TRUE))>1){
    ordAbund<-rowSums(dat[,colnames(dat) %in% name])
    return(ordAbund)
  }else{
    return(dat[,name])
  }
}, dat=orders))
colnames(ordAbund)<-levels(as.factor(names$order))

ordAbund<-cbind(combdat[,c("Season","SeagrassCover","TowArea")],ordAbund)
```
 
```{r}
ordAbund<-write.csv(ordAbund,"cleanedCommunityOrders.csv",quote=FALSE,row.names = FALSE)
```

 
# Pipefish data

Get the size data from the google doc

```{r googleDrive, eval=FALSE}
library("googledrive")
drive_dat<-drive_download("NZsyngnathids_measurements", 
                          path="../data/NZsyngnathids_measurements",
                          type = "xlsx",
                          overwrite=TRUE)
```
```{r GetXLSXData}
library(openxlsx)
sTraits<-read.xlsx("../data/NZsyngnathids_measurements.xlsx",sheet = "S.nigra",
                                       skipEmptyRows = TRUE,detectDates = TRUE)
lTraits<-read.xlsx("../data/NZsyngnathids_measurements.xlsx",sheet = "L.elevatus",
                                       skipEmptyRows = TRUE,detectDates = TRUE)


infoCols<-1:11
datCols<-c(grep("TotalLength",colnames(sTraits)),
           grep("SVL",colnames(sTraits)),
           grep("BodyWidth",colnames(sTraits)),
           grep("BodyDepth",colnames(sTraits)),
           grep("SnoutLength",colnames(sTraits)),
          grep("SnoutWidth",colnames(sTraits)),
          grep("SnoutDepth",colnames(sTraits)),
          grep("HeadLength",colnames(sTraits)))
pipeTraits<-dplyr::bind_rows(sTraits[,c(infoCols,datCols)],lTraits[,c(datCols,infoCols)])
pipeTraits<-pipeTraits[pipeTraits$LocationName=="Duvauchelle Bay",]
```
```{r AvgVals}
pipeTraits$TotalLength_Avg<-rowMeans(pipeTraits[,grep("TotalLength",colnames(pipeTraits))],na.rm = TRUE)
pipeTraits$SVL_Avg<-rowMeans(pipeTraits[,grep("SVL",colnames(pipeTraits))],na.rm = TRUE)
pipeTraits$BodyWidth_Avg<-rowMeans(pipeTraits[,grep("BodyWidth",colnames(pipeTraits))],na.rm = TRUE)
pipeTraits$BodyDepth_Avg<-rowMeans(pipeTraits[,grep("BodyDepth",colnames(pipeTraits))],na.rm = TRUE)
pipeTraits$SnoutLength_Avg<-rowMeans(pipeTraits[,grep("SnoutLength",colnames(pipeTraits))],na.rm = TRUE)
pipeTraits$SnoutWidth_Avg<-rowMeans(pipeTraits[,c(grep("SnoutWidth",colnames(pipeTraits)),grep("SnoutDepth",colnames(pipeTraits)))],na.rm = TRUE)
pipeTraits$HeadLength_Avg<-rowMeans(pipeTraits[,grep("HeadLength",colnames(pipeTraits))],na.rm = TRUE)
```
```{r pipeTraitsKeepCols}
pipeTraits<-pipeTraits[!is.na(pipeTraits$Species),c(colnames(pipeTraits)[1:10],grep("Avg",colnames(pipeTraits),value=TRUE))]
```
```{r pipeTraitsFactors}
pipeTraits$Season<-gsub("(\\d{4})\\-(\\d{2})\\-(\\d{2})","\\2",pipeTraits$DateSampled)
pipeTraits$Season[pipeTraits$Season %in% c("09","10","11")]<-"Spring"
pipeTraits$Season[pipeTraits$Season %in% c("12","01","02")]<-"Summer"
pipeTraits$Season[pipeTraits$Season %in% c("03","04","05")]<-"Fall"
pipeTraits$Season[pipeTraits$Season %in% c("06","07","08")]<-"Winter"
pipeTraits$Season<-factor(pipeTraits$Season,levels=c("Winter","Spring","Summer","Fall"))
# fill in the missing measurements
pipeTraits$TotalLength_Avg[is.na(pipeTraits$TotalLength_Avg)]<-pipeTraits$Length_field[is.na(pipeTraits$TotalLength_Avg)]
# fill in status
pipeTraits$Status[pipeTraits$Sex=="female"]<-"female"
pipeTraits$Status<-as.factor(pipeTraits$Status)
```
```{r savePipeTraits, eval=FALSE}
write.csv(pipeTraits,"../data/pipefishTraits.csv",row.names = FALSE,quote=FALSE)
```


Combine counts and trait data -- it is not perfect. 

I was hoping to correct the pipefish data to be adult/juvenile, if possible.

```{r pipeLong}
pipe_long<-tidyr::gather(pipedat, size, count, 10:13, factor_key=TRUE)
pipe_long$species<-gsub("(.*)_.*$","\\1",pipe_long$size)
pipe_long$species[pipe_long$species=="s.nigra"]<-"stigmatopora"

pipe_long$size<-factor(gsub("(.*)_","",pipe_long$size),
                           levels = c("below50mm","above50mm"))
pipe_long$density<-pipe_long$count/pipe_long$towArea
```

```{r pipefishCnts}
set.seed(51435)
pipefishCnts<-pipe_long
pipefishCnts$size<-as.character(pipefishCnts$size)



for( date in unique(pipedat$Day)){
  trt<-pipeTraits[pipeTraits$DateSampled == date,]
  cnt<-pipe_long[pipe_long$Day == date,]
  
  st<-trt[trt$Species=="Stigmatopora nigra" ,]
  sc<-cnt[cnt$species=="stigmatopora" & cnt$count>0,]
  
  if(length(sc$size[sc$size %in% "above50mm"]) == length(st$Sex[st$Sex %in% c("male","female")])){
    pipefishCnts$size[pipefishCnts$Day==date &
                        pipefishCnts$species=="stigmatopora" &
                        pipefishCnts$count>0 &
                        pipefishCnts$size=="above50mm"]<-"adult"
  }else if(length(sc$size[sc$size %in% "above50mm"]) > length(st$Sex[st$Sex %in% c("male","female")])){
    
    # if there are more 'above' ones than adults, randomly choose one to be the adult
    # and make the others juveniles
    abv<-row.names(pipefishCnts[which(pipefishCnts$Day==date &
                 pipefishCnts$species=="stigmatopora" &
                 pipefishCnts$count>0 &
                 pipefishCnts$size=="above50mm"),])
    adt<-sample(abv,length(st$Sex[st$Sex %in% c("male","female")]))
    juv<-abv[! abv %in% adt]
    print(paste(length(juv),">50mm Stigmatopora reassigned as juveniles;", length(adt), ">50mm inds retained as adults"))
    pipefishCnts[adt,"size"]<-"adult"
    pipefishCnts[juv,"size"]<-"juvenile"
  } 
  if(length(pipefishCnts$size[pipefishCnts$Day==date &
                        pipefishCnts$species=="stigmatopora" &
                        pipefishCnts$count>0 &
                        pipefishCnts$size=="above50mm"]) > 0){
    browser()
  }
  
  st<-trt[trt$Species=="Leptonotus elevatus" ,]
  sc<-cnt[cnt$species=="leptonotus" & cnt$count>0,]
  
  if(length(sc$size[sc$size %in% "above50mm"]) == length(st$Sex[st$Sex %in% c("male","female")])){
    pipefishCnts$size[pipefishCnts$Day==date &
                        pipefishCnts$species=="leptonotus" &
                        pipefishCnts$count>0 &
                        pipefishCnts$size=="above50mm"]<-"adult"
  }else if(length(sc$size[sc$size %in% "above50mm"]) > length(st$Sex[st$Sex %in% c("male","female")])){
    
    # if there are more 'above' ones than adults, randomly choose one to be the adult
    # and make the others juveniles
    abv<-which(pipefishCnts$Day==date &
                 pipefishCnts$species=="leptonotus" &
                 pipefishCnts$count>0 &
                 pipefishCnts$size=="above50mm")
    
    adt<-sample(abv,length(st$Sex[st$Sex %in% c("male","female")]))
    juv<-abv[! abv %in% adt]
    print(paste(length(juv),">50mm Leptonotus reassigned as juveniles;", length(adt), ">50mm inds retained as adults"))
    pipefishCnts$size[adt]<-"adult"
    pipefishCnts$size[adt]<-"juvenile"
  } else{
    browser()
  }
  
}

pipefishCnts$size[pipefishCnts$size=="below50mm"]<-"juvenile"
pipefishCnts$size[pipefishCnts$count==0]<-"absent"
```
```{r savePipefishCnts, eval=FALSE}
write.csv(pipefishCnts,"../data/pipefishCnts.csv",row.names = FALSE,quote=FALSE)
```


# Other size-classed organisms

```{r}
grep("mm$",colnames(rawdat),value=TRUE)
```

