---
title: "Data processing"
author: "Sarah Flanagan"
date: "16/06/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "./figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="./figs/")
```

```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(dplyr)
library(tidyr)
library(geosphere)
```

# Community data

```{r}
rawdat<-read.csv("../data/MonthlyRawSampling_Duvauchelle_2020.csv")

# remove unnecessary column
rawdat<-subset(rawdat,select=-X)

```

First, check for NAs that should be zeros

```{r}
rawdat[,17:62][is.na(rawdat[,17:62])]<-0
```

Then add a date column

```{r}
rawdat$Day<-as.Date(
  paste0(
    gsub("^(\\d)$","0\\1",rawdat$Date),
    gsub("^(\\d)$","0\\1",rawdat$Month),
    rawdat$Year),
  format = "%d%m%Y")
```

And group them by season

```{r}
rawdat$Season<-gsub("\\d+-(\\d+)-\\d+","\\1",rawdat$Day)
rawdat$Season[rawdat$Season %in% c("09","10","11")]<-"Spring"
rawdat$Season[rawdat$Season %in% c("12","01","02")]<-"Summer"
rawdat$Season[rawdat$Season %in% c("03","04","05")]<-"Fall"
rawdat$Season[rawdat$Season %in% c("06","07","08")]<-"Winter"
```


Calculate the distance of the tows

```{r}
towDist<-apply(rawdat,1,function(dat){
  distm(c(as.numeric(dat["long_start"]),
          as.numeric(dat["lat_start"])),
        c(as.numeric(dat["long_stop"]),
          as.numeric(dat["lat_stop"])))
})
# turn zeroes into 1s
towDist[which(towDist==0)]<-1

# add to data sheet
rawdat$towDist<-towDist
```

Finally remove any rows with missing seagrass covers

```{r}
rawdat<-rawdat[!is.na(rawdat$seagrass_cover),]
```

Fix inconsistent seagrass cover typing

```{r}
rawdat$seagrass_cover<-tolower(rawdat$seagrass_cover)
rawdat$seagrass_cover[rawdat$seagrass_cover=="spare"]<-"sparse"
```

Combine columns so that each organism has its a single column

```{r}
combNames<-grep("mm$",colnames(rawdat),value=TRUE)
combNames<-combNames[combNames!="whitebait_below50mm"] # only whitebait instance
combPatts<-unique(gsub("_.*$","",combNames))
combPatts[combPatts=="austrovenus"]<-"venus" # one included 'venus_bivalve'

totals<-as.data.frame(do.call(cbind,lapply(combPatts,function(pattern,dat){
  tot<-rowSums(dat[,grep(pattern,colnames(dat))])
  return(tot)
},dat=rawdat)))
colnames(totals)<-combPatts
colnames(totals)[colnames(totals)=="venus"]<-"austrovenus"
totals$isopod<-rowSums(rawdat[,c("isopod","green_centipede")])

nonCombNames<-colnames(rawdat[,17:62])[!colnames(rawdat[,17:62]) %in% combNames]
nonCombNames<-grep("wrasse",nonCombNames,invert=TRUE,value=TRUE)
nonCombNames<-grep("venus",nonCombNames,invert=TRUE,value=TRUE)
nonCombNames<-nonCombNames[!nonCombNames %in% c("isopod","green_centipede")]
```



Create summary data frame

```{r}
combdat<-data.frame(cbind(
  Date=rawdat$Day,
  Season=rawdat$Season,
  TowNum=rownames(rawdat),
  TowDist=rawdat$towDist,
  SeagrassCover=rawdat$seagrass_cover,
  totals,
  rawdat[,nonCombNames]
  )
)
```

```{r}
write.csv(combDat,"cleanedCommunity.csv",quote=FALSE, row.names = FALSE)
```


I wonder if I need to group these by a common taxonomic level (Family or Order, most likely)

Shrimp species likely included in our counts:

- *Palaemon affinis* (glass shrimp), family Palaemonidae, infraorder Caridea
- *Alphaeus novazealandiae* or *A. socialis* (snapping shrimp | kowhitiwhiti moana), family alpheidae, infraorder caridea
- *Alope spinifrons* (painted shrimp), family Hippolytidae, infraorder Caridea
- *Pontophilus* or *Parapontophilus* species, infraorder Caridea
- *Hippolyte* spp. (chamaeleon shrimp), family Hippolytidae, infraorder Caridea

```{r}
names<-read.csv("../data/species_names.csv")
kable(names,booktabs=TRUE)
```

Should I combine any of these, and if so, at what level?
 
 
# Pipefish data

```{r}
lengths<-read.csv("../data/pipefish.lengths.csv")
counts<-read.csv("../data/pipefish.counts.csv")
```


# Other size-classed organisms

```{r}
grep("mm$",colnames(rawdat),value=TRUE)
```

