---
title: "Testing hierarchical pairwise beta"
author: "Sarah Flanagan"
date: "19/07/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.extra='',fig.pos="H",
                      fig.path = "../figs/",dpi = 300,fig.keep='last',dev='png')
knitr::opts_knit$set(root.dir='../results/',fig_path="../figs/")
```
```{r loadLibs, include=FALSE, echo=FALSE}
library(scales)
library(knitr)
library(kableExtra)
library(vegan)
library(dplyr)
library(tidyr)

files.sources <- list.files(path = "../R/hierDiversity",full.names = TRUE)
tmp<-sapply(files.sources, source)
source("../R/PairwiseBetaFunctions.R")

tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
set.seed(8675309) # Jenny Jenny
```


I've added the calculation of pairwise turnover to the hierDiversity R code. I did so by incorporating Zach's functions from the script `diversityFunctionsNGBA.R` (which was associated with the pairwise beta paper) into the `div.part` and `dz` functions in hierDiversity. But now I want to test to make sure that the function is behaving as expected, so I'll test it using the simulations/data Zach used for the pairwise beta paper (from `PairwiseBetaSource.R` downloaded as an Appendix). 

He used the BCI dataset, and did a bunch of sub-samples to test the method. I think I can just compare his estimates of 'true' diversity to estimates using hierDiversity.

```{r loadBCI}
### Load BCI dataset from vegan package
data(BCI) 
bci <- as.matrix(BCI)
```

```{r PairwiseCalcs, eval=FALSE}
### We were only interested in diversity orders 0, 1, & 2 
### but this vector could contain any order of interest.
q.seq <- 0:2 

##### BCI actual abundances  
trueDiv <- noHier(comm = bci, q.seq)
trueTurn <- turnover(trueDiv$betaT, N=nrow(bci))
trueLocal <- mapply(local,q=q.seq, beta=trueDiv$betaT, N=nrow(bci))
trueReg <- mapply(regional,q=q.seq, beta=trueDiv$betaT, N=nrow(bci))
truePWTurn <- trueDiv$pBeta
truePWLocal <- trueDiv$lOver
truePWRegional <- trueDiv$rOver

notPWTrue <- rbind(trueTurn, trueLocal, trueReg)
PWTrue <- rbind(truePWTurn, truePWLocal, truePWRegional)


saveRDS(trueDiv,"trueDiv.RDS")
```
```{r}
trueDiv<-readRDS("trueDiv.RDS")
```


```{r hierDiversity, eval=FALSE}
# least inclusive to most inclusive L->R
grps<-as.matrix(cbind(rownames(bci),"Site"))

hierResults<-lapply(q.seq,
                    hierDiversity,
                    dat=bci, 
                    group=grps, 
                    replace = NULL,
                    reps = 99, 
                    quant = c(0.025, 0.975), 
                    sims = FALSE)
saveRDS(hierResults,"trueHier.RDS")
```
```{r}
hierResults<-readRDS("trueHier.RDS")
hierResultsTab<-list(hierResults[[1]]$lev2$Site,
                     hierResults[[2]]$lev2$Site,
                     hierResults[[3]]$lev2$Site)
```

```







